---
author: "CC"
date: "2023-03-27"
output:
    html_document:
        theme:
            bootswatch: yeti
        toc: yes
        toc_float:
          collapsed: false
          smooth_scroll: true
        number_sections: yes
        df_print: kable
        code_folding: hide
    pdf_document:
        number_sections: yes
        toc: yes
        toc_depth: 3
        keep_tex: no
title: |
  | `r DATASET`
  | Protocole 2: Processing
---

<!-- Javascript for zooming on figures (adapted from: https://stackoverflow.com/questions/40401680) -->

<!-- Jquery import conflicts with DT::datatable so needs to be commented here -->
<!-- <script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"></script> -->

<style>
.zoomDiv {
  display: none;
  position: fixed;
  top: 50%;
  left: 50%;
  z-index: 50;
  transform: translate(-50%, -50%);
  background-color: #FFFFFF;
  box-shadow: 0px 0px 50px #888888;
  width: fit-content;
  max-width: 90%;
  max-height: 90%;
  overflow: auto;
}

.zoomImg {
  width: 150%;
}
</style>

<script type = "text/javascript">
  $(document).ready(function() {
    $("body").prepend("<div class = \"zoomDiv\"><img src = \"\" class = \"zoomImg\"></div>");
    // onClick for all img except the zoomed one and link ones (filter)
    // use "img.zoom" and out.extra = "class = \"zoom\"" in chunk to specify manually which chunk images can be zoomed
    $("img:not(.zoomImg)").filter(":not(a *)").click(function() {
      $(".zoomImg").attr("src", $(this).attr("src"));
      $(".zoomDiv").show();
    })
    // onClick function for hiding div
    $("img.zoomImg").click(function() {
      $(".zoomDiv").hide();
    })
  })
</script>


```{r setup, include = FALSE}
options(knitr.purl.inline = TRUE)
knitr::opts_chunk$set(
  # code evaluation
  eval = TRUE,

  # text output
  echo = TRUE,
  results = "hold",
  warning = FALSE,
  error = FALSE,
  message = FALSE,
  strip.white = TRUE,

  # code decoration
  #tidy = TRUE,
  tidy.opts = list(width.cutoff = 90),
  comment = "",
  attr.output = ".numberLines",

  # plots
  #fig.path = "figure/",      # is set later, in chunk setup2
  fig.show = "asis",         # tuned to "hold" in multiple plots chunk
  dev = c("png", "pdf"),
  fig.width = 12,
  fig.height = 12,
  #fig.asp = 1.3               # fig.height = fig.width * fig.asp
  #out.extra = "style = "border:5px solid orange""    # orange box arround plots
  fig.align = "center"   # should be tuned to default in multiple plots chunk
)
```

```{r load-libraries, include = FALSE}
library(kableExtra)
library(dplyr)
library(Seurat)
library(stringr)
library(ggplot2)
```

```{r get-params, include = TRUE}
# bindingIsLocked("params", env = .GlobalEnv)
# unlockBinding("params", env = .GlobalEnv)
# attach(params)
set.seed(general_seed)
```

```{r dir-managment, include = TRUE}
if (!dir.exists(PATH_ROOT)) { dir.create(PATH_ROOT) }
if (!dir.exists(PATH_OUT_HTML)) { dir.create(PATH_OUT_HTML) }
PATH_OUT_ANALYSIS <- file.path(PATH_ROOT, "04_process")
if (!dir.exists(PATH_OUT_ANALYSIS)) { dir.create(PATH_OUT_ANALYSIS) }
if (!dir.exists(PATH_OUT_FIG)) { dir.create(PATH_OUT_FIG, recursive = TRUE) }
```

<!-- # Global settings -->

```{r show-params, eval=FALSE, include=TRUE}
ul <- unlist(params)
df <- data.frame(keyName = names(ul), value = ul, row.names = NULL)
df <- df[! grepl("^PATH|DATASET", df$keyName), ]
names(df) <- c("Parameters", "Values")
df1 <- df[1:ceiling(dim(df)[1] / 3), ]
df1$Parameters <- cell_spec(df1$Parameters, bold = TRUE)
df2 <- df[(ceiling(dim(df)[1] / 3) + 1) : (2 * ceiling(dim(df)[1] / 3)), ]
df2$Parameters <- cell_spec(df2$Parameters, bold = TRUE)
df3 <- df[((2 * ceiling(dim(df)[1] / 3)) + 1):dim(df)[1], ]
df3$Parameters <- cell_spec(df3$Parameters, bold = TRUE)


list(df1, df2, df3) %>%
  knitr::kable(align = "ll", row.names = FALSE, escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE)
rm(df)
```

```{r setup2, eval = TRUE, include =TRUE}
knitr::opts_chunk$set(
  # code evaluation
  eval = TRUE,

  # figures
  fig.path = paste0(PATH_OUT_FIG, "/02_process_")
)
```

# Load `r DATASET` data

```{r load-data}
SO <- readRDS(paste0(PATH_RDS_OBJECTS, "/01_qc_", DATASET, "_", FILTER, ".rds"))
```

# Preprocessing workflow

```{r hvg-vst2, eval=hvg_meth != "mvp", echo=hvg_meth != "mvp", message=TRUE}
SO <- NormalizeData(SO, normalization.method = norm_meth, verbose=TRUE)
SO <- ScaleData(SO, features=rownames(SO), do.scale=do_scale, verbose=TRUE)
saveRDS(SO, file=file.path(PATH_RDS_OBJECTS, paste0("02_normANDscaled_", DATASET, "_", FILTER, ".rds")))

SO <- FindVariableFeatures(SO, nfeatures=hvg_num, selection.method = hvg_meth, verbose=TRUE)
var.feats.dataset <- VariableFeatures(SO)

SO <- RunPCA(SO, npcs = pca_npcs, nfeatures.print = pca_print, seed.use = general_seed, verbose=TRUE)
SO <- RunUMAP(SO, dims = 1:top_pcs, seed.use = general_seed, verbose = TRUE)
```

```{r hvg-mvp2, eval=hvg_meth == "mvp", echo=hvg_meth == "mvp", message=TRUE}
t(as.data.frame(table(SO@assays[["RNA"]]@meta.features$mvp.variable))) %>%
  knitr::kable(caption = "Number of features selected as highly variable genes (HVG)",
               row.names = TRUE,
               escape = F,
               align = "cc") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

var.feats.DATASET <- VariableFeatures(SO)
#saveRDS(var.feats.DATASET, file=file.path(PATH_RDS_OBJECTS, paste0("var.feats.dataset_", DATASET, ".rds")))

SO <- ScaleData(SO, features=rownames(SO), do.scale=do_scale, verbose=TRUE)
SO <- RunPCA(SO, npcs = pca_npcs, nfeatures.print = pca_print, seed.use = general_seed, verbose=TRUE)
SO <- RunUMAP(SO, dims = 1:top_pcs, seed.use = general_seed, verbose = TRUE)
```

```{r preprocessSinglet, out.width='50%', results='asis'}
#resolutions <- c(seq(0.1, 0.4, 0.1), seq(0.6, 1.6, 0.2))
resolutions <- clust_res
SO <- FindNeighbors(SO, dims = 1:top_pcs, verbose=TRUE)
SO <- FindClusters(
    SO,
    resolution = resolutions,
    algorithm = clust_meth,
    random.seed = general_seed,
    verbose=TRUE
)
```


```{r dimplot-cluster-cplt, eval=FILTER == "complete", echo=FILTER == "complete", out.width="50%"}
Idents(SO) <- SO$seurat_clusters
DimPlot(SO)
```

```{r dimplot-cluster-fltd, eval=FILTER == "filtered", echo=FILTER == "filtered", out.width="50%"}
Idents(SO) <- SO$seurat_clusters
DimPlot(SO)
```

# Exports

```{r save4}
# variable features
write.table(
    var.feats.dataset,
    file = file.path(
        PATH_OUT_ANALYSIS,
        paste0("varFeatures_meth.", hvg_meth, "_nb.", hvg_num, "_", DATASET, "_", FILTER, ".txt")),
    row.names = TRUE,
    col.names = TRUE
)

# PCA and UMAP tables
invisible(sapply(c("pca", "umap"), function(dr) {
    SO@reductions[[dr]]@cell.embeddings <- cbind(
        cellIDs = rownames(SO@reductions[[dr]]@cell.embeddings),
        SO@reductions[[dr]]@cell.embeddings
    )
    write.table(
        SO@reductions[[dr]]@cell.embeddings,
        file = file.path(
            PATH_OUT_ANALYSIS,
            paste0("DR_", dr, "_", DATASET, "_", FILTER, ".csv")),
        row.names = TRUE,
        col.names = TRUE,
        sep = ","
    )
}))

# Clustering
keepMDnames <- c("orig.ident", "nCount_RNA", "nFeature_RNA")
SO@meta.data <- cbind(cellIDs = rownames(SO@meta.data), SO@meta.data)
exportMD <- names(SO@meta.data)[! names(SO@meta.data) %in% keepMDnames]

write.table(
    SO@meta.data[exportMD],
    file.path(
        PATH_OUT_ANALYSIS, 
        paste0("clustering_", DATASET, "_", FILTER, "_meth.", clust_meth, "_res.", clust_res, ".csv")),
    row.names = TRUE,
    col.names = TRUE,
    sep = ","
)

invisible(gc())
```






