---
title: "gastD11"
author: "BULAVKA"
date: "2023-05-17"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(Seurat)
library(patchwork)
library(kableExtra)
library(stringr)
library(DoubletFinder)
source("/mnt/DATA_4TB/projects/gastruloids_sc_Lescroart/analysis/embryE9gastD10/02_scripts/extract_topn.R")
source("/mnt/DATA_4TB/projects/gastruloids_sc_Lescroart/analysis/embryE9gastD10/02_scripts/add_dominantMetadata.R")

```
# 1 Load the dataset 
```{r Load the dataset, include=FALSE}
gs.data <- Read10X(data.dir = "/mnt/DATA_4TB/projects/gastruloids_sc_Lescroart/analysis/embryE9gastD10/00_rawData/gastD11",  gene = 2)
gs <- CreateSeuratObject(counts = gs.data, project = "gastD11")
# gs.brut <-  CreateSeuratObject(counts = gs.data, project = "gastD11")
```
# 2. Parametres
```{r Parametres, echo=FALSE}
## rds & tables
rdsObject <- "/mnt/DATA_4TB/projects/gastruloids_sc_Lescroart/analysis/embryE9gastD10/50_rdsObjects/"
## table 
table.path <- "/mnt/DATA_4TB/projects/gastruloids_sc_Lescroart/analysis/embryE9gastD10/53_OutPutTable"
## QC
mito <- 7.5
ribo <- 10
min_counts <- 4000
max_counts <- 90000
min_feats <- 150
min.cells <- 3
## Normalization 
norm_method <- "LogNormalize"
##  Identification of highly variable features (feature selection)
var_feat_method <- "vst"
## Clustering
Leiden <- 4 
res_var <- c(1, 1.5, 2, 3)
general.seed <- 42
## FindDoublets
dblt.rate <-8
colors.table <- read.table(file="/mnt/DATA_4TB/projects/gastruloids_sc_Lescroart/analysis/embryE9gastD10/InputTables/ClusterColors.tsv", sep="\t", header=T, comment.char="", as.is=T)
colors.celltype <- setNames(colors.table$blind_friendly[!is.na(colors.table$transferred_identity)], colors.table$transferred_identity[!is.na(colors.table$transferred_identity)])

# Genes of interest
gene_of_interest <- read.csv("/mnt/DATA_4TB/projects/gastruloids_sc_Lescroart/analysis/embryE9gastD10/genesOfInterest.tsv", header = TRUE)

```
# 3. QC 
```{r QC, echo=FALSE, message=FALSE, warning=FALSE}
df <- data.frame(rowSums(gs@assays$RNA@counts != 0))
df$features <- rownames(df)
colnames(df) <- c('Nbr_of_cells', 'features')
rownames(df) <- NULL
df_non0 <- df[df$Nbr_of_cells > 0, ]
gs[["percent.mt"]] <- PercentageFeatureSet(gs, pattern = "^mt-")
gs[["percent.ribo"]] <- PercentageFeatureSet(gs, pattern="^Rp[sl]")  
gs <- subset(gs, features=which(df$Nbr_of_cells > min.cells), subset = nCount_RNA >= min_counts & nCount_RNA <= max_counts
             & nFeature_RNA >= min_feats & percent.mt <= mito & percent.ribo > ribo)

##  Visualize QC metrics as a violin plot  
VlnPlot(gs, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.ribo"),  ncol = 4) 
## FeatureScatter
plot1<- FeatureScatter(gs, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") +geom_smooth (method= 'lm')
plot2<- FeatureScatter(gs, feature1 = "nCount_RNA", feature2 = "percent.mt") +geom_smooth (method= 'lm')
plot3<- FeatureScatter(gs, feature1 = "nCount_RNA", feature2 = "percent.ribo") +geom_smooth (method= 'lm')
plot4<- FeatureScatter(gs, feature1 = "percent.mt", feature2 = "percent.ribo") +geom_smooth (method= 'lm')
plot1 + plot2 + plot3 + plot4
```

# 4. Normalization and Identification of highly variable features (feature selection) 
```{r Normalization and Identification of highly variable features, message=FALSE, warning=FALSE}
gs <- NormalizeData(gs, normalization.method = norm_method, scale.factor = 10000, verbose = FALSE)
gs <- FindVariableFeatures(gs, selection.method = var_feat_method, nfeatures = 2000, verbose = FALSE)
## Identify the 10 most highly variable genes-----------------------------------
top20 <- head(VariableFeatures(gs), 20)
## plot variable features with and without labels ------------------------------
plot1 <- VariableFeaturePlot(gs)
plot2 <- LabelPoints(plot = plot1, points = top20, repel = TRUE)
plot2

```

# 5. Scaling the data and perform linear dimensional reduction 
```{r Scaling the data and perform linear dimensional reduction, message=FALSE, warning=FALSE}
all.genes <- rownames(gs) 
gs <- ScaleData(gs, features = all.genes, verbose = FALSE)
gs <- RunPCA(gs, features = VariableFeatures(object = gs), npcs = 20, nfeatures.print = 10, verbose=TRUE, seed.use=general.seed)
ElbowPlot(gs, ndims = 20)
gs <- FindNeighbors(gs, dims = 1:20, verbose=FALSE)
gs <- FindClusters(gs, resolution = res_var, algorithm = Leiden, random.seed = general.seed) # algorithm = algo.cluster, random.seed = general.seed, verbose=FALSE
gs <- RunUMAP(gs, dims = 1:20, seed.use = general.seed, verbose = FALSE)
DimPlot(gs, group.by = "RNA_snn_res.1" , reduction = "umap", label = TRUE) + NoLegend()


```

# 6. Finding differentially expressed features (cluster biomarkers)
```{r Finding differentially expressed features (cluster biomarkers), message=FALSE, warning=FALSE}
gs.markers <- FindAllMarkers(gs, group.by = "RNA_snn_res.1", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, verbose=FALSE)
gs.markers %>%
  group_by(cluster) %>%
  top_n(n = 5, wt = avg_log2FC) -> top10
DoHeatmap(gs,  group.by = "RNA_snn_res.1", features = top10$gene) + NoLegend()
```
# 7. Doubets Finder
```{r Doubets Finder, message=FALSE, warning=FALSE, results='hide'}
atlas.subset <- readRDS("/mnt/DATA_4TB/projects/gastruloids_sc_Lescroart/analysis/embryE9gastD10/50_rdsObjects/atlas_preprocessed2.rds")
anchors <- FindTransferAnchors(reference = atlas.subset, query = gs,
                               dims = 1:20, reference.reduction = "pca")
predictions <- TransferData(anchorset = anchors, refdata = atlas.subset$celltype, verbose = FALSE,
                            dims = 1:20)
gs <- AddMetaData(gs, metadata = predictions[,"predicted.id"], col.name = "celltype_DF")

#
Idents(gs) <- gs@meta.data$celltype_DF

p1 <- DimPlot(gs,
        pt.size = 1,
        repel = TRUE,
        label = TRUE,
        label.size = 6,
        cols = colors.celltype[levels(Idents(gs))]) +
  ggtitle( " Cell identities to perform DoubletFinder") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text = element_blank(),
        line = element_blank()) +
  NoLegend()
p1

```

## pK identifikation and Homotypic Doublets Proportion Estimate
```{r pK identifikation and Homotypic Doublets Proportion Estimate, message=FALSE, warning=FALSE, results='hide'}
## pK identifikation
sweep.res <- paramSweep_v3(gs, PCs = 1:20) # as estimated from PC elbowPlot
sweep.stats_gs <- summarizeSweep(sweep.res, GT= FALSE)
Sys.sleep(0.5)
bcmvn_gs <- find.pK(sweep.stats_gs)

ggplot(bcmvn_gs, aes(pK, BCmetric, group = 1)) +
  geom_point() +
  geom_line()

pK <- bcmvn_gs %>% # select the pK that corresponds to max bcmvn to optimize doublet detection
  filter(BCmetric == max(BCmetric)) %>%
  select(pK) 

pK <- as.numeric(as.character(pK[[1]]))

## Homotypic Doublets Proportion Estimate 
annotations <- gs@meta.data$celltype_DF
homotypic.prop <- modelHomotypic(annotations)
nDoublets <- round(ncol(gs)*dblt.rate/100) # poi
nDoublets_nonhomo <- round(nDoublets*(1-homotypic.prop)) #poi.adj
## run doubletFinder: creating artificial doublets 
gs <- doubletFinder_v3(gs,
                       PCs = 1:20,
                       pN = 0.25,
                       pK = pK,
                       nExp = nDoublets_nonhomo)

## Plot the singlets, doublets and the count of UMIs in each cells
col_dblts <- grep("DF.classifications", colnames(gs@meta.data), value=TRUE)

Idents(gs) <- col_dblts
cellsData <- data.frame(gs@reductions[["umap"]]@cell.embeddings, gs@meta.data[,col_dblts])
colnames(cellsData) <- c(colnames(gs@reductions[["umap"]]@cell.embeddings), "col_dblts")
p1 <- ggplot(cellsData[c('UMAP_1', 'UMAP_2')], # Omit the column used for facetting to get all points repeated in all facets
             aes( x = UMAP_1,
                  y = UMAP_2)) +
  geom_point( alpha = .4,
              size  = .75,
              color = "grey") +
  geom_point( data = cellsData, # Now provide data including column for facetting
              aes(color = col_dblts),
              alpha = .5,
              size  = 1.2) +
  facet_wrap(facets = vars(factor(col_dblts, levels = c("Singlet", "Doublet"))),
             ncol = 2) +
  NoLegend() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text = element_blank(),
        line = element_blank())


p1

cellsData <- data.frame(gs@reductions[["umap"]]@cell.embeddings, gs@meta.data$nCount_RNA)
colnames(cellsData) <- c(colnames(gs@reductions[["umap"]]@cell.embeddings), "nCount_RNA")
Db <- table(gs@meta.data[col_dblts])["Doublet"]
saveRDS(gs, file = paste0(rdsObject, "gast.rds"))


```
# 8. Preprocessing workflow (after doublets rgsoval)
```{r Preprocessing workflow (after doublets rgsoval), message=FALSE, warning=FALSE}
gs.sing <- subset(gs, idents='Singlet')
saveRDS(gs.sing, file = paste0(rdsObject, "gast_singlets.rds"))
gs.sing <- FindVariableFeatures(gs.sing, selection.method = var_feat_method, nfeatures=2000, verbose=FALSE)
gs.sing <- ScaleData(gs.sing, features=rownames(gs.sing), do.scale=FALSE, verbose=FALSE)
gs.sing <- RunPCA(gs.sing, npcs = 20, nfeatures.print = 10, seed.use = general.seed, verbose=TRUE)
# Clustering 
gs.sing <- FindNeighbors(gs.sing, dims = 1:20, verbose=FALSE)
gs.sing <- FindClusters(gs.sing, resolution = res_var, algorithm = Leiden, random.seed = general.seed, verbose = FALSE) # random.seed = general.seed, verbose=FALSE
## Run non-linear dimensional reduction (UMAP/tSNE) 
gs.sing <- RunUMAP(gs.sing, dims = 1:20, seed.use = general.seed, verbose = FALSE)
DimPlot(gs.sing, group.by = "RNA_snn_res.1" , reduction = "umap", label = TRUE) + NoLegend()
# Finding differentially expressed features (cluster biomarkers)------------
gs.sing.markers <- FindAllMarkers(gs.sing, group.by = "RNA_snn_res.1" , only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, verbose = FALSE)
top10 <- gs.sing.markers %>%
  group_by(cluster) %>%
  top_n(n = 5, wt = avg_log2FC) 
DoHeatmap(gs.sing, features = top10$gene, group.by = "RNA_snn_res.1") + NoLegend()
write.table(gs.sing.markers, file = paste0(table.path, "marquers_gastruloides_total.csv"))


```
# 9. Cell identities after perform DoubletFinder
```{r Cell identities after perform DoubletFinder, message=FALSE, warning=FALSE}
anchors <- FindTransferAnchors(reference = atlas.subset, query = gs.sing,
                               dims = 1:20, reference.reduction = "pca")
predictions <- TransferData(anchorset = anchors, refdata = atlas.subset$celltype,
                            dims = 1:20)
gs.sing <- AddMetaData(gs.sing, metadata = predictions[,"predicted.id"], col.name = "celltype_sing")
Idents(gs.sing) <- gs.sing@meta.data$celltype_sing

DimPlot(gs.sing,
        pt.size = 1,
        repel = TRUE,
        label = TRUE,
        label.size = 6,
        cols = colors.celltype[levels(Idents(gs.sing))]) +
  ggtitle( "Cell identities") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text = element_blank(),
        line = element_blank()) +
  NoLegend()
saveRDS(gs.sing, file = paste0(rdsObject,"gast_sing_process.rds"))


```
# Expression of Flt1 and Flt1-ps
```{r Expression of Flt1 and Flt1-ps, message=FALSE, warning=FALSE}

FeaturePlot(gs.sing, features = c("Ftl1","Ftl1-ps1"),  reduction = "umap", cols = c("lightgrey", "darkred"), ncol = 3) & theme(plot.title = element_text(size = 10))

saveRDS(gs.sing, file = paste0(rdsObject,"gs_sing_total.rds"))

```

